Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-05-22 WSSSPE2_report/WSSSPE2_review.tex
Update WSSSPE2_review.tex

\todo{short intro to group here}
Six papers were categorized under the theme of Exploring Sustainability. The group included four of the authors from the six papers submitted and an additional number of participants who expressed an interest in the theme at the workshop. Each paper had a different perspective on the concept of sustainability, which ranged from the sustainment of communities to defining sustainability as a first-class, composite non-functional requirement.
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Discussion and Actions}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
The papers that were discussed in the Exploring Sustainability group are:
\begin{itemize}
\item Mario Rosado de Souza, Robert Haines, and Caroline Jay. Defining
Sustainability through Developers' Eyes: Recommendations from an Interview
Study~\cite{wssspe2_rosada_de_souza}

>>>>

\todo{short intro to group here}
Six papers were categorized under the theme of Exploring Sustainability. The group included four of the authors from the six papers submitted and an additional number of participants who expressed an interest in the theme at the workshop. Each paper had a different perspective on the concept of sustainability, which ranged from the sustainment of communities to defining sustainability as a first-class, composite non-functional requirement.
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Discussion and Actions}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsubsection{Discussion}

Each author was invited to outline the key action from their paper as a potential discussion point for the group; in their absence, the group facilitators outlined the actionable points from papers where the authors were not present. The key actions from the six papers included recommendations for improving practice in sustainable software engineering ~\cite{wssspe2_rosada_de_souza, wssspe2_downs}, the development of Software and Infrastructure as a Service as a mechanism for fostering sustainable science communities ~\cite{wssspe2_patra}, developer incentives for code contributions to open source projects~\cite{wssspe2_pierce}, the establishment of a set software engineering principles based in scalability, reproducibility, and energy efficiency ~\cite{wssspe2_shi}, and applying software architectures as a mechanism for architectural-level reasoning about sustainability ~\cite{wssspe2_venters}.

The group took a position to view sustainability from the perspective of addressing the challenges related to the development, deployment, and maintenance of reusable software~cite{}.

The principal focus of the initial discussion considered how to foster cultural change towards developing sustainable software in academic environments. It was suggested that a new requirement driven by the agencies funding research projects where software was an intrinsic part of enabling the research programme would be an additional element within the grant proposal to provide a sustainability plan for sustaining the software. This type of initiative would provide the necessary incentive and motivation for researchers to consider how to sustain their software beyond the lifetime of the project.
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
The papers that were discussed in the Exploring Sustainability group are:
\begin{itemize}
\item Mario Rosado de Souza, Robert Haines, and Caroline Jay. Defining
Sustainability through Developers' Eyes: Recommendations from an Interview
Study~\cite{wssspe2_rosada_de_souza}
